---
title: "Math 295: Homework 1"
subtitle: "Introduction to Data Visualisation"
author: "Your Name Here"  
date: "Due Tuesday, January 24, 2023"
output: 
  github_document: 
    fig_height: 4
    fig_width: 6
    toc: true
    toc_depth: 2
---

```{r setup, include = FALSE}
# Don't change this setup chunk. 
# It sets global defaults and includes the tidyverse packages
# The option include = FALSE prevents warnings and messages from printing to your report.
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Overview

Using the `diamonds` data set contained in the `ggplot2` package, this homework asks you to examine factors that affect the price of a diamond.  Because the setup chunk above contains the option `echo=TRUE`, whenever you knit your R Markdown file, your document will contain your R commands, the R output and  your narrative text. Submit by pushing to your GitHub repository. You are required to have at least 3 commits. I recommend that the first be after you have added your name to this file above. 

# About the Data Set


1) Use R help to learn about the `diamonds` data set.  Answer the following questions:  
    * How many cases are there in the total data set?  
    * How many variables are there?  
    * What are the units used for the `price` variable?  
    * What are the possible values of the `color` variable? Which is best?  

**Answer:**  

```{r}

```

# Creating the Subset for Analysis

Note: As you should have seen above, the `diamonds` data set is very large so for the rest of the homework you will work with a randomly selected subset of 2000 cases created by the following code. You MUST set `eval=TRUE` in the following code to create the subset. Run this chunk by selecting the green arrow at the top of the chunk.^[The `set.seed()` command initializes a sequence of pseudo random numbers so the sampling is reproducible. That is, I will get the same random numbers as you did when I knit your file.]

```{r, diamonddata, eval=FALSE}
# Change eval=FALSE to eval=TRUE when you knit this or it won't have any effect
set.seed(2950000) # Replace the 4 zeroes at the end with your birthday MMDD as in 0526 for May 26
diamond_sub <- diamonds %>% 
  filter(carat <= 2.5) %>%
  sample_n(2000)
```

# Analysis of the `diamond_sub` dataset  

**Note**: For all of the following, be sure to use appropriate `ggplot2` procedures. Use of other R graphics platforms will not earn credit.  


2) Create a scatterplot of diamond price (`y`) versus carat weight (`x`).  Comment briefly on the nature of the relationship.  

**Answer:**  

```{r}

```



3) Modify your plot in 2) setting the `alpha` transparency value to 1/10. Why is this useful for large datasets?  

**Answer:**  

```{r}

```


4) Modify your plot in 2), mapping the `clarity` of the diamond to the color aesthetic of the point. Comment on what this plot says about the relationship between `clarity` and the two other variables?  

**Answer:**  

```{r}

```



5) Once again, modify your plot in 2) mapping the `clarity` of the diamond to the shape aesthetic of the point. Explain why you get warning messages for this plot.     

**Answer:**  

```{r}

```


6) Which of the two plots above do you prefer, the one that maps `clarity` to color or shape?  Justify your choice.  

**Answer:**  



7) Examine the relationship between `price` and `carat` weight in more detail  by creating one plot for each diamond  `color`.  Describe what you can learn from this analysis.  

**Answer:** 

```{r}

```


8) Just for fun, map diamond `color` to the aesthetic color so that the points on each plot are a different color.  

**Answer:** 

```{r}

```



9) Extend your plot in 7) to obtain a separate scatterplot of the relationship between `price` and `carat` weight for every combination of diamond `cut` and `color`. (Hint: You should have 35 individual scatterplots.)

**Answer:** 

```{r}

```



10) Challenge: Create a brand new scatterplot of the relationship between `price` and `carat` weight but in this plot use aesthetics to color all points correponding to diamonds with depth (`z`) greater than 3.5mm a different color than those with smaller depths. What is striking about this plot?  

**Answer:**

```{r}

```

### Note